\documentclass[oneside, 10pt]{amsart}
\usepackage{amscd, amsmath, amssymb, amsthm, amsfonts, amstext, verbatim, mathtools, xfrac, microtype, nameref, thmtools}
\usepackage[breaklinks=true]{hyperref}
\usepackage[hyperref=true, backend=bibtex, firstinits=true, citestyle=numeric-comp, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}
\usepackage[capitalize]{cleveref}
\usepackage[matrix,arrow,curve]{xy}
\usepackage{enumitem}

\addbibresource{paper.bib}
\renewbibmacro*{volume+number+eid}{\ifentrytype{article}{\- \iffieldundef{volume}{}{Vol.~\printfield{volume},}\iffieldundef{number}{}{ No.~\printfield{number},}}}
\renewbibmacro{in:}{\ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}}
\newbibmacro{string+doi}[1]{\iffieldundef{doi}{\iffieldundef{url}{#1}{\href{\thefield{url}}{#1}}}{\href{http://dx.doi.org/\thefield{doi}}{#1}}}
\DeclareFieldFormat[article, inproceedings, inbook, book, thesis]{title}{\usebibmacro{string+doi}{\mkbibquote{#1}}}
\renewcommand*{\bibfont}{\footnotesize}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\T}{T}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\rk}{rk}

\newcommand{\rA}{\mathsf{A}}
\newcommand{\rB}{\mathsf{B}}
\newcommand{\rC}{\mathsf{C}}
\newcommand{\rD}{\mathsf{D}}
\newcommand{\rE}{\mathsf{E}}
\newcommand{\rF}{\mathsf{F}}
\newcommand{\rG}{\mathsf{G}}

\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\XX}[1]{\mathcal{X}_{#1}}
\newcommand{\RR}[1]{\mathcal{R}_{#1}}

\newtheorem{thm}{Theorem}
\Crefname{thm}{Theorem}{Theorems}
\numberwithin{equation}{section}

\newtheorem{lemma}{Lemma}
\numberwithin{lemma}{section}
\Crefname{lemma}{Lemma}{Lemmas}
\newlist{lemlist}{enumerate}{1} \setlist[lemlist]{label={\rm(\arabic{lemlisti})}, ref=\thelemma.(\arabic{lemlisti}),noitemsep} \Crefname{lemlisti}{Lemma}{Lemma}

\newtheorem{cor}[lemma]{Corollary}
\Crefname{cor}{Corollary}{Corollaries}

\newtheorem{prop}[lemma]{Proposition}
\Crefname{prop}{Proposition}{Propositions}

\newtheorem*{thm*}{Theorem}
\newtheorem*{lemma*}{Lemma}

\theoremstyle{definition}

\newtheorem{dfn}[lemma]{Definition}
\Crefname{dfn}{Definition}{Definitions}
\newtheorem{example}[lemma]{Example}
\Crefname{example}{Example}{Examples}

\theoremstyle{remark}

\newtheorem{rem}[lemma]{Remark}
\Crefname{rem}{Remark}{Remarks}


\title{On $K_2$-analogue of Serre problem for $\SL_4$}
\keywords {Steinberg group, $K_2$-functor, Serre problem
 {\em Mathematical Subject Classification (2010):} 19C20}
\author {Sergey Sinchuk}
\address{Chebyshev laboratory, St. Petersburg State University, St. Petersburg, Russia}
\email {sinchukss at gmail.com}
\date {\today}

\begin{document}   
\maketitle

\section{Introduction}

\section{Preliminaries}

Our notation and conventions follows~\cite[\S~4]{Vav09}.
Let $\Phi$ be an irreducible root system with some fixed basis of simple roots $\Pi = \{\alpha_1, \ldots, \alpha_\ell\}$.
For a root $\alpha\in\Phi$ we denote by $m_i(\alpha)$ the $i$-th coefficient in the expansion of $\alpha$ in $\Pi$,
 i.\,e. $\alpha = \sum_{i=1}^n m_i(\alpha) \alpha_i$.

We denote by $\Phi^\vee$ the \emph{dual root system of $\Phi$} consisting of coroots $\alpha^\vee = 2\alpha/(\alpha, \alpha)$, $\alpha\in \Phi$.
As usual, $P(\Phi^\vee)$ denotes the lattice spanned by the \emph{fundamental weights $\varpi_i$},
 that are uniquely determined by relations $\langle\varpi_i, \alpha_j^\vee \rangle = (\varpi_i, \alpha_j) = \delta_{ij}.$ %TODO: Mistake?

\subsection{Steinberg groups and relative Steinberg groups}

\subsection{Rehmann-Soule theorem}

\section{Overview of Tulenbayev's proof of $K_2$-analogue of Serre problem}

\section{Proof of the main result}

\subsection{$\sigma$-pairs}
Notice that for $\varpi \in P(\Phi^\vee)$ and $\beta \in \ZZ \Phi$ one has $(\varpi, \beta) \in \ZZ $.
Consequently, for $\varepsilon \in R^*$ and $\varpi \in P(\Phi^\vee)$ the formula $\chi_{\varpi, \varepsilon}(\beta) = \varepsilon ^ {(\varpi, \beta)}$
defines a character $\chi_{\varpi, \varepsilon} \in \Hom(\ZZ \Phi, R^*)$.

Consider the action of $H=\Hom(\ZZ \Phi, R^*)$ on the set of generators $\mathcal{X}_{\Phi, R}$ of the Steinberg group $\St(\Phi, R)$ given by
\begin{equation} \chi \cdot x_\alpha(\xi) = x_\alpha(\chi(\alpha) \cdot \xi),\ \chi \in H,\ \alpha\in \Phi,\ \xi \in R. \end{equation}
Since $\chi$ is a character, this action respects the set of Steinberg relations $\mathcal{R}_{\Phi, R}$ and therefore gives a well-defined action of $H$ on $\St(\Phi, R)$.

In the sequel we will be mostly interested in the following example.
\begin{example}
Let $A$ be a ring, denote by $R$ the ring $A[t, t^{-1}]$ of Laurent polynomials over $A$ and let $\alpha_i \in \Pi$ be a simple root.
Define automorphisms $\beta_i^+$ and $\beta_i^-$ of $\St(\Phi, R)$ via $\beta_i^+ = \chi_{\varpi_i, t}$, $\beta_i^- = \chi_{\varpi_i, t^{-1}}$.
It is easy to see that
\begin{equation}\label{eq:sigma_act} \beta_i^\pm(x_\alpha(\xi)) = x_\alpha(t^{\pm m_i(\alpha)} \cdot \xi).\end{equation}
\end{example}

One of the key steps of in the proof of Suslin lemma for $K_2$ is to define analogue of maps $\beta_i^\pm$ on the level of the group $\St(\Phi, A[t])$.
Of course, one cannot expect such maps to be automorphisms or even be defined on the whole group $\St(\Phi, A[t])$.
Thus, our first goal is to define the subgroups of $\St(\Phi, A[t])$ on which these morphisms could, in principle, be defined.

In the sequel we assume that $i$ is a vertex of the Dynkin diagram $D(\Phi)$ of $\Phi$ with the property that $m_i(\alpha_{\max})=1$ (of course, this implies that $\Phi$ is not of type $\rE_8, \rF_4$ or $\rG_2$).
Denote by $\mathcal{X}_i$ (resp. $\mathcal{X}^i$) the subsets of $\mathcal{X}_{\Phi, A[t]}$ consisting of those generators $x_{\alpha}(\xi)$ of $\St(\Phi, A[t])$ for which 
$m_i(\alpha) > 0$ (resp. $m_i(\alpha)<0$) implies that $\xi \in A[t]$ is divisible by $t$.
% either $i\neq k$ (resp. $j\neq k$) or $\xi\in R[t]$ is divisible by $t$.
Denote by $N_i$ and $N^i$ the subgroups of $\St(\Phi, A[t])$ generated by $\mathcal{X}_i$ and $\mathcal{X}^i$, respectively.
%Our goal is to construct mutually inverse morphisms $\sigma_k$, $\sigma_k^{-1}$ fitting into the following diagram.

\begin{dfn} A pair of mutually inverse maps $(\sigma_i^+, \sigma_i^-)$ between the groups $N^i$ and $N_i$ will be called a {\it $\sigma$-pair corresponding to the vertex $i\in D(\Phi)$},
 if these maps fit into the following diagram.
\begin{equation} \label{eq:sigma-diagram}
\xymatrix{ N_i\ar@{-->}[r]^{\sigma_i^+}\ar[d] & N^i \ar[d] \ar@{-->}[r]^{\sigma_i^-} & N_i \ar[d] \\ 
          \St(\Phi, R) \ar@<-0.0ex>[r]_{\beta_i^+} & \St(\Phi, R) \ar@<-0.0ex>[r]_{\beta_i^-} & \St(\Phi, R)} 
\end{equation}
In the above diagram the vertical arrows are canonical embeddings of $N_k$ and $N^k$ into $\St(\Phi, A[t])$ postcomposed with 
 the map of Steinberg groups induced by the embedding $A[t] \hookrightarrow R$.
%As before, we denote by $R$ the ring $A[t, t^{-1}]$ of Laurent polynomials over $A$.
         % \GL(n, R) \ar@<-0.0ex>[r]_{(-)^{\delta_k}} & \GL(n, R) \ar@<-0.0ex>[r]_{{}^{\delta_k}(-)} & \GL(n, R)
\end{dfn}

For the remainder of this subsection our main goal will be to verify that an analogue of the group decomposition from \cite[Lemma~3.1f]{Tu83} holds for $\St(\Phi, A)$
 provided $\St(\Phi, R)$ admits at least one $\sigma$-pair.

Denote by $\widetilde{W}$ the subgroup of $\St(\Phi, A)$ generated by elements $w_\alpha(1)$.
The subgroup $\widetilde{W}$ acts on $\St(\Phi, A)$ via conjugation, for example, 
 the action on the generators of $\mathcal{X}_{\Phi, A}$ can be expressed via the formula ${}^{w_\alpha(1)} \!x_{\beta}(\xi) = x_{\sigma_{\alpha}(\beta)} ( \pm \xi)$.
It is easy to see that the action of $\widetilde{W}$ on the set of root subgroups $\{X_\alpha \mid \alpha \in \Phi \}$ of $\St(\Phi, R)$ coincides with the obvious
 action of the Weyl group $W(\Phi)$ on it. Since $N_i$ and $N^i$ are generated by root subgroups we immediately obtain the following statement.
\begin{lemma} The set of elements in the orbit of $N_i$ (resp. $N^i$) under the conjugation action of $\widetilde{W}$ 
is in bijective correspondence with 
the set of adjacency classes of $W(\Phi)$ with respect to its Levi subgroup $W_i = W(\Delta_i)$. %TODO: Unsure whether it is true at all :)
\end{lemma}
In the sequel we denote by $\Sigma$ the set of elements in $\widetilde{W}$ whose projections in $W(\Phi)$ form the set of representatives for
 right adjacency classes of $W(\Phi)$ with respect to $W_i$. %We can assume these representatives to be short


\subsection{Structure theorem for Steinberg groups}


sufficient conditions implying the main structure result for Steinberg groups.

Let $\mu$ be a microweight of the root system $\Phi$. Denote by $N_\mu$

\subsection{Construction of \texorpdfstring{$\sigma$}{sigma}-morphisms for $\SL_4$}
Let $R$ be a local ring and assume that $n\geq 4$.
For $1\leq k\leq n$ denote by $\delta_k$ the $n\times n$ matrix of $\GL(n, R[t, t^{-1}])$ which differs from the unit matrix 
 only by the element $t$ on the $k$-th place of its diagonal.
Recall from~\cite[Corollary~4]{Ka77} that there exists unique automorphisms $\beta_{\delta_k}$, $\beta_{\delta_k^{-1}}$ of $\St(n, R[t, t^{-1}])$
 modeling conjugation with matrices $\delta_k$ and $\delta_k^{-1}$ (see diagram~\eqref{eq:sigma-diagram} below)

For $n\geq 5$ $\sigma-$morphisms have already been constructed by Tulenbaev, see \S~3 of~\cite{Tu83}.
Thus, in the sequel we are mainly interested in the case $n=4$.
To simplify notation, we limit ourselves to constructing the map $\sigma_1$, so everywhere in the sequel we assume that $k=1$. 

Of course, the map of right conjugation with $\delta_1$ induces the obvious one-to-one set-theoretic correspondence between $\mathcal{X}_1$ and $\mathcal{X}^1$. 
However, there seems to be no reasonable description of the complete set of relations between the elements $\mathcal{X}_1$ that would allow one to describe $N_1$
 as a group given by generators and relations.
Thus, in order to define $\sigma_1$ we will use the alternate description of $N^1$ given in terms of ``another presentation''.
We start with the following simple observation.
\begin{lemma} \label{lem:n1-decomp} There is an isomorphism $N_1 \cong N_{1,0} \rtimes P_1^-(R)$, where $N_{1,0}$ denotes the subgroup $\St(n, R[t], tR[t])$
 and $P_1^-(R)$ is the subgroup of $\St(n, R)$ generated by $x_{ij}(\xi)$ with $i\neq 1$.
\end{lemma}

Recall that the universal property of semidirect products asserts that for any group $H$ acting on a group $N$ via $\phi \colon H \to \Aut(N)$ and any group homomorphisms
 $f_N\colon N \to G$, $f_H\colon H \to G$ satisfying 
\begin{equation} \label{eq:coherence-condition} f_N(\phi(h)(n)) = f_H(h) f_N(n) f_H(h)^{-1},\ n\in N,\ h\in H,\end{equation} there exists a unique map $f\colon N \rtimes H \to G$
 extending $f_N$ and $f_H$. 
Thus, by the above lemma, in order to construct the map $\sigma_1$ we need to construct the following two maps and verify that they satisfy~\eqref{eq:coherence-condition}:
\[ (\sigma_1)_{P_1^-(R)} \colon P_1^-(R) \to N^1, \ \ (\sigma_1)_{N_{1,0}} \colon N_{1,0} \to N^1.\]

It is easy to define the first map $(\sigma_1)_{P_1^-(R)}$.
Using the decomposition $P_1^-(R) \cong U^-_1 \rtimes L_1$ where 
\[U^-_1 = \langle x_{i1}(\xi) \mid i\neq 1,\ \xi\in R \rangle \text{ and } L_1 = \langle x_{ij}(\xi) \mid i,  j \neq 1,\ \xi\in R\rangle \]
we apply the universal property of semidirect products once again 
 and define $(\sigma_1)_{P_1^-(R)}$ by requiring that it acts identically on $L_1$ 
and acts on elements of $U^-_1$ via the formula $(\sigma_1)_{P_1^-(R)}(x_{i1}(\xi))= x_{i1}(t\xi)$.

Before we proceed with the construction of the second map $(\sigma_1)_{N_{1,0}}$ we need to recall several technical
 constructions pertaining to ``another presentation'' of the group $N_{1,0}$.

For $u \in R^n$ we denote by $D(u)$ the subset of $R^n$ consisting of all $v$ satisfying $u^tv = 0$ and having at least two zero entries.
Recall from 3.2 of~\cite{Ka77} that for every $u, v, w \in R[t]^n$ such that $u^t \cdot v = 0$ there
 is a decomposition of $(w^t u) \cdot v$ into a sum of elements of $D(u)$, called {\it canonical decomposition}:
\setcounter{equation}{1}
\renewcommand{\theequation}{\arabic{equation}}
\begin{equation} \label{eq:canonical} (w^tu) \cdot v=\sum_{i<j}u_{ij} c_{ij}(v, w),\end{equation}
where $u_{ij}=e_iu_j-e_ju_i,$ and $c_{ij}(v, w)=v_iw_j-v_jw_i.$

Let $A$ be a commutative ring and $I$ be a splitting ideal of $A$.
Let $u, v \in A^n$ be such that $u^t v = 0$ and assume, moreover, that either $u$ or $v$ has at least one zero entry.
Recall from 3.8, 3.10 of~\cite{Ka77} that under these assumptions one can define certain element
 $x(u, v)$ of $\St(n, A)$ such that $\pi(x(u, v)) = t(u, v) = 1 + uv^t$.
The following properties of $x(u, v)$ are standard (cf.~\cite[Lemma~1.1]{Tu83}).
\begin{lemma} \label{lem:xsmall-properties}
\begin{lemlist}
 \item \label{item:xsmall-scalar} If $v$ or $w$ has at least two zero components, then 
 \begin{equation}\nonumber x(v, wa) = x(va, w).\end{equation}
 \item \label{item:xsmall-additivity} If $w_1$ and $w_2$ have at least one common zero coordinate then
 \begin{equation}\nonumber x(v, w_1)x(v, w_2) = x(v, w_1+w_2).\end{equation}
 \item \label{item:xsmall-commute} If $v$, $v'$ are simultaneously orthogonal to $w$ and $w'$ and the elements $w, w'$
  both have at least two zero components then $x(v, w)$ and $x(v', w')$ commute.
\end{lemlist}
\end{lemma}

Let $u \in E(n, A)e_1$, $v \in I^n$ be such that $u^tv = 0$.
Recall from \S~4 of~\cite{LS17} that one can define the following elements of $\St(n, A, I)$:
\begin{equation} \label{eq:sigma-definition}
 F(u, v) = \prod\limits_{i=1}^r x(u,  v_i),\ \
 S(v, u) = \prod\limits_{i=1}^r x( v_i, u).
\end{equation}
Here $\{v_r\}$, $r=1,\ldots,N$ is any collection of elements of $I^n \cap D(u)$ such that $\sum_{r=1}^N v_r = v$.
Since $u$ is unimodular, such a collection always exists by~\eqref{eq:canonical}, moreover
 $F(u, v)$ and $S(u, v)$ do not depend on its choice. %TODO: Give reference

Recall from~\cite{LS17} that it is possible to give a presentation of the relative Steinberg group $\St(n, A, I)$ formulated in terms of elements $F(u, v)$, $S(v, u)$.
More formally, \cite[Proposition 3.10]{LS17} asserts that $\St(n,\,A,\,I)$ admits presentation with the set of generators
\begin{multline*} \{F(u,\,v)\mid u\in E(n,\,A)e_1,\ v\in I^n,\ u^tv=0\}\, \cup \\ \cup \{S(u,\,v)\mid u\in I^n,\ v\in E(n,\,A)e_1,\ u^tv=0\}\end{multline*}
subject to relations
\setcounter{equation}{0}
\renewcommand{\theequation}{R\arabic{equation}}

\begin{align}
&F(u,\,v)F(u,\,w)=F(u,\,v+w), \label{add4}\\
&S(u,\,v)S(w,\,v)=S(u+w,\,v), \label{add5}\\
&F(u,\,v)F(u',\,v')F(u,\,v)^{-1}=F(t(u,\,v)u',\,t(v,\,u)^{-1} v'), \label{conj3} \\
&F(ge_1,\,g^*e_2a)=S(ge_1a,\,g^*e_2),\ \text{for all}\ a\in I,\, g \in E(n, A).
\end{align}
Here $g^*$ denotes the contragradient matrix, i.\,e. $g^* = {g^{-1}}^t$ and $e_i$ is the standard basis vector of $A^n$.

\setcounter{equation}{3}
\renewcommand{\theequation}{\arabic{equation}}

To define the map $(\sigma_1)_{N_{1,0}}$ we will invoke the above presentation in the special case $A=R[t]$, $I=tR[t]$.
Indeed, define $\sigma_1$ on $N_{1,0}=\St(n, R[t], tR[t])$ by the following identities: 
\begin{equation}
 \sigma_1(F(u, v)) = \prod\limits_{i=1}^r x(\delta_1^{-1} \cdot t u, \delta_1 \cdot t^{-1} v_i),\ \
 \sigma_1(S(v, u)) = \prod\limits_{i=1}^r x(\delta_1^{-1} \cdot v_i, \delta_1 \cdot u).
\end{equation}

\begin{lemma} \label{lem:cor-conj}
 \begin{lemlist}
   \item \label{item:correctness} The elements $\sigma_1(F(u, v))$ and $\sigma_1(S(v, u))$ are well-defined, i.\,e. they do not depend on the choice of decomposition for $v$.
   \item \label{item:conjugation} For any $g \in N^1$ one has 
      \begin{equation}
          \nonumber {}^g\left(\sigma_1(F(u, v))\right) = \sigma_1\left(F(g' \cdot u, {g'}^{*} \cdot v)\right),\ \ 
                    {}^g\left(\sigma_1(S(u, v))\right) = \sigma_1\left(S(g' \cdot u, {g'}^{*} \cdot v)\right),                    
      \end{equation}
      where $g' = {}^{\delta_1}\pi(g).$
 \end{lemlist}
\end{lemma}
\begin{proof}[Short proof]
 The first assertion can be easily verified using the canonical decomposition and~\cref{lem:xsmall-properties}
  (compare with Tulenbaev's argument after~\cite[Lemma~1.1]{Tu83}).
  
 The second assertion can be proved similarly to~\cite[3.14]{Ka77}, (cf. also with the proof of~\cite[Lemma~4.4d]{LS17}).
\end{proof}
\begin{proof}[Long proof]
 Let us verify the first assertion. Let $v = \sum_r v^r$ be a decomposition as above. 
 Since each $v^r$ is orthogonal to $u$ we can write the canonical decomposition  
 $v^r = \sum_{i<j} u_{ij} c_{ij}(v^r, w)$, moreover, $\sum_{r} c_{ij}(v^r, w) = c_{ij}(v, w)$.
 \cref{lem:xsmall-properties} allows us to rewrite the right-hand side of~\ref{eq:sigma-definition} in such a way that 
  it does not depend on the choice of decomposition for $v$:
 \begin{multline}
  \prod\limits_r x(\delta_1^{-1} ut, \delta_1 v^rt^{-1}) = \prod\limits_{r}\prod\limits_{i<j} x(\delta_1^{-1} ut, \delta_1 u_{ij}c_{ij}(v^r, w)t^{-1}) = \\ 
   = \prod\limits_{i<j} x(\delta_1^{-1} ut, \delta_1 u_{ij}c_{ij}(v, w)t^{-1}). 
 \end{multline}

 We only verify the second assertion for $F(u, v)$  and only in the special case $g = x_{hk}(\xi) \in \mathcal{X}^1$, $g' = t_{hk}(\xi') \in \pi(\mathcal{X}_1)$.
 Choose $w\in R^n$ so that $w^t u = 1$ and write $v = \sum_{1\leq i<j\leq n} u_{ij} c_{ij}$, where $u_{ij}$ and $c_{ij} = c_{ij}(v, w)$ are as in~\eqref{eq:canonical}. 
 We can write ${}^g\left(\sigma_1\left(F(u, v)\right)\right)$ as follows:
 \begin{equation} \nonumber
   \prod\limits_{1\leq i<j\leq n} x(\pi(g) \cdot \delta_1^{-1} \cdot t u, \pi(g)^* \cdot \delta_1 \cdot t^{-1} u_{ij}c_{ij})
                                   = \prod\limits_{1\leq i<j\leq n} x(\delta_1^{-1} \cdot g'\cdot t u, \delta_1 \cdot {g'}^* \cdot t^{-1} u_{ij}c_{ij}).
 \end{equation} 
 Now for every factor in the right-hand side we do the following.
 If $\{i, j\} = \{h, k\}$ or $\{i, j\} \cap \{h, k\} = \emptyset$ we leave the term unchanged.
 Otherwise, if, $|\{i, j\} \cap \{h, k\}| = 1$,  say $j = h$, $i\neq k$, we further decompose it as follows:
  \begin{equation} \nonumber
    x(\delta_1^{-1} \cdot t u', \delta_1 \cdot {g'}^* \cdot u_{ij} (c_{ij} t^{-1})) = 
     x(\delta_1^{-1} \cdot u' t, \delta_1 \cdot u'_{ij} (c_{ij} t^{-1})) \cdot 
     x(\delta_1^{-1} \cdot u' t, \delta_1 \cdot u'_{ki} (c_{ij} t^{-1})).
  \end{equation}
  Here $u' = g'  u = t_{jk}(\xi') \cdot u$.
  Notice that ${g'}^*u_{ij} = t_{kj}(-\xi') \cdot u_{ij} = e_iu_j - e_ju_i + e_k\xi'u_i =u'_{ij}+u'_{ki}\xi'$. 
  
  Thus, we arrive at a product $\prod_{s} x(\delta_1^{-1} \cdot u't, \delta_1 \cdot v'_s t^{-1}) $, where $v'_s \in D(u')$ and
   $\sum v'_s = {g'}^*v$. Clearly, this product equals $\sigma_1\left(F(g' \cdot u, {g'}^{*} \cdot v)\right)$ by definition.
\end{proof}

\begin{cor}
 The map $\sigma_1$ preserves relations~\eqref{add4}--\eqref{conj3}.
\end{cor}
\begin{proof}
 For~\eqref{add4}--\eqref{add5} this is an immediate corollary of the first assertion of the previous lemma.
 For~\eqref{conj3} this follows from the second part of the lemma applied in the special case $g = \sigma_1(F(u, v)) \in N^1$.
\end{proof}

It remains to verify that $\sigma_1$ preserves~\eqref{coinc}.
Let us do this at first in the special case when $g^*$ belongs to the subset $G_0 = H_{12}(R) \cdot U^+_1(R) \cdot U^-_1(R)$.
Here we denote by $H_{12}(R)$ the subgroup of $T(n, R)$ generated by semisimple root elements $ h_{12}(\xi)$, $\xi \in R^*$.

Since the only nonzero components of $u = g^* \cdot e_2$ are $u_1$ and $u_2$ we can present $v = g \cdot e_1$ as a sum $v' + v''$, where 
 $v'_i=0$, $i>2$ and $v''_1=v''_2=0$.
Since $v', v'' \in D(u)$ and $u \in D(v)$, from the definition of $\sigma_1$ we obtain that:
\begin{multline} \label{eq:special-case}
 \sigma_1(S(g \cdot e_1a, g^* \cdot e_2)) = x(\delta_1^{-1} \cdot v'a, \delta_1\cdot  u) \cdot x(\delta_1^{-1}\cdot v''a, \delta_1 \cdot u) = \\
  = x(\delta_1^{-1} va, \delta_1 \cdot u) = x(\delta_1^{-1} vat^{-1}, \delta_1 \cdot ut) = \sigma_1(F(g \cdot e_1, g^* \cdot e_2a)) 
\end{multline}

To obtain the assertion in the general case notice that for $R$ local the group $\E(n, R)$ admits the following decomposition:
\[\E(n, R) = \mathrm{EP}_1(R) \cdot H_{12}(R) \cdot U^-_1(R) \cdot U^+_1(R).\]
This can be either obtained as a corollary of Gauss decomposition or can be proved by a direct calculation. %TODO: Give reference

Applying transpose-inverse automorphism $(-)^*$ to the above decomposition and invoking~\cref{lem:n1-decomp} we obtain that $\E(n, R[t]) = \pi(N_1) \cdot G_0$.
Thus, every $g \in \E(n, R[t])$ can be factored as $\pi(n) \cdot h$ for some $n\in N_1$ and $h \in G_0$.
Since $\pi(n) = {}^{\delta_1} \pi(n')$ for some $n' \in N^1$ it remains to apply~\cref{item:conjugation} and~\eqref{eq:special-case} to obtain the required assertion:
\begin{multline} \nonumber \sigma_1(F(g \cdot e_1, g^* \cdot e_2a)) = {}^{n'}(\sigma_1(F(h \cdot e_1, h^* \cdot e_2a))) = \\
                           = {}^{n'}(\sigma_1(S(h \cdot e_1a, h^* \cdot e_2))) = \sigma_1(S(g \cdot e_1a, g^* \cdot e_2)). \end{multline}

Thus, we have completed the construction of the map $(\sigma_1)_{N_{1,0}}$.
It is not hard to verify that the condition~\eqref{eq:coherence-condition} is satisfied with these definitions,
 therefore, we have also completed the construction of the map $\sigma_1$.

\subsection{Proof of the theorem on \texorpdfstring{$\mathbb{P}^1$}{P1}-glueing for $K_2$}

\section{Applications}

\subsection{Bruhat--Tits buildings}

\subsection{\texorpdfstring{$A^1$}{A1}-invariance of the fundamental group}

\DeclareRobustCommand{\VAN}[2]{#2}
\printbibliography

\end{document}
